---
title: "Analysis of navigation tree metrics"
output: html_document
---

```{r load data}
DATADIR <- sprintf('%s/wikimedia/trunk/data', Sys.getenv('HOME'))
PLOTDIR <- sprintf('%s/snap-papers/2015/west1-ashwinp-wmf/FIG', Sys.getenv('HOME'))
data <- read.table(pipe(sprintf('gunzip -c %s/tree_metrics_sample_en.tsv.gz', DATADIR)), header=TRUE)

# Set this to FALSE if you don't want to save plots to PDFs.
save_plots <- TRUE

.default_par <- par(no.readonly=TRUE)
```

We analyze a sample of 300,000 navigation traces.
This data was computed before we considered searches and other 302 queries.

# Tree size

Tree size is defined as the number of nodes in the tree.

First, a histogram (log-log scales).

```{r}

if (save_plots) pdf(sprintf('%s/tree_properties.pdf', PLOTDIR), width=1.68, height=1.5, pointsize=6,
                    family='Helvetica', useDingbats=FALSE)
par(mar=c(3.4, 3.6, 1.1, 0.6))
plot(sort(data$size), rev(1:sum(is.finite(data$size)))/sum(is.finite(data$size)), log='xy', type='l', bty='n',
     xlim=c(1,200), ylim=c(1e-5,1), xlab='', ylab='', col=rgb(.9,.6,0))
lines(sort(data$degree_mean), rev(1:sum(is.finite(data$degree_mean)))/sum(is.finite(data$degree_mean)))
mtext(expression(italic(n)), side=1, line=2.4)
mtext(expression(Pr(italic(X) >= italic(n))), side=2, line=2.4)
legend('topright', legend=c(expression(paste(italic(X), ' = tree size')), expression(paste(italic(X), ' = avg. node deg.'))),
       col=c(rgb(.9,.6,0), 'black'), lty=1, bty='n')
par(.default_par)
if (save_plots) dev.off()

# Fraction of non-singleton trees with avg. degree >= 3.
mean(data$degree_mean >= 3, na.rm=TRUE)
# Absolute number.
sum(data$degree_mean >= 3, na.rm=TRUE)/nrow(data)*3e9

metric <- data$size
# PDF.
h <- hist(metric, breaks=(min(metric)-1):max(metric), plot=FALSE)
plot(h$breaks[-1], h$counts/sum(h$counts), log='xy', panel.first=grid(),
     xlab='Tree size', ylab='Probability')
# CCDF.
if (save_plots) pdf(sprintf('%s/tree_metrics_size.pdf', PLOTDIR), width=1.68, height=1.2, pointsize=6,
                    family='Helvetica', useDingbats=FALSE)
par(mar=c(3.4, 3.4, 1.1, 0.6))
plot(sort(metric), rev(1:sum(is.finite(metric)))/sum(is.finite(metric)), log='xy', type='l',
     xlim=c(1,200), ylim=c(1e-5,1), panel.first=grid(), xlab='', ylab='')
mtext('n', side=1, line=2.4)
mtext(expression(paste('Pr[num nodes >= n]')), side=2, line=2.4)
par(.default_par)
if (save_plots) dev.off()
```

If the distribution were exponential, $\frac{1}{1-F(s)}$ would be a straight line on a single-log plot (where $s$ is the tree size, and $F(s)$, the CDF), but clearly this isn't the case:

```{r}
x <- sort(data$size)
y <- (1:length(data$size))/length(data$size)
plot(x, 1/(1-y), log='y', ylim=c(1,1e5), panel.first=grid(), xlab='Tree size s', ylab='1/(1-F(s))')
```

If the distribution were Weibull, $\log(\frac{1}{1-F(s)})$ would be a straight line on log-log scales (left).
If the distribution were a power law, $1-F(s)$ would be a straight line on log-log scales (right).
(see http://www.itl.nist.gov/div898/handbook/apr/section2/apr221.htm)

```{r}
par(mfrow=c(1,2))
plot(x, log(1/(1-y)), log='xy', ylim=c(0.5,15), panel.first=grid(), xlab='Tree size s', ylab='log(1/(1-F(s)))')
plot(x, rev(y), log='xy', panel.first=grid(), xlab='Tree size s', ylab='1 - F(s)')
```

TODO: Figure out the best fit via a goodness-of-fit or likelihood ratio test.

# Tree depth

Note that depth cannot be larger than 100, since we impose this limit at data-extraction time (to avoid stack overflows in recursive tree processing).

```{r}
metric <- data$depth
h <- hist(metric, breaks=(min(metric)-1):max(metric), plot=FALSE)
plot(h$breaks[-1], h$counts/sum(h$counts), log='xy', panel.first=grid(),
     xlab='Tree depth', ylab='Probability')
# CCDF.
if (save_plots) pdf(sprintf('%s/tree_metrics_depth.pdf', PLOTDIR), width=1.68, height=1.2, pointsize=6,
                    family='Helvetica', useDingbats=FALSE)
par(mar=c(3.4, 3.4, 1.1, 0.6))
plot(sort(metric), rev(1:sum(is.finite(metric)))/sum(is.finite(metric)), log='xy', type='l',
     xlim=c(1,200), ylim=c(1e-5,1),
     panel.first=grid(), xlab='', ylab='')
mtext('d', side=1, line=2.4)
mtext(expression(paste('Pr[depth >= d]')), side=2, line=2.4)
par(.default_par)
if (save_plots) dev.off()
```

Again, if the distribution were Weibull, the left plot would be a straight line; if it were a power law, the right plot would be a straight line.

```{r}
x <- sort(data$depth)
y <- (1:length(data$depth))/length(data$depth)
par(mfrow=c(1,2))
plot(x, log(1/(1-y)), log='xy', panel.first=grid(), ylim=c(1,15), xlab='Tree size s', ylab='log(1/(1-F(s)))')
plot(x, rev(y), log='xy', panel.first=grid(), xlab='Tree size s', ylab='1-F(s)')
```

# Node degree

Maximum degree per tree.

```{r}
# PDF
metric <- data$degree_max
h <- hist(metric, breaks=(floor(min(metric, na.rm=TRUE)-1)):ceiling(max(metric, na.rm=TRUE)), plot=FALSE)
plot(h$breaks[-1], h$counts/sum(h$counts), log='xy', panel.first=grid(),
     xlab='Max non-leaf degree', ylab='Probability')
# CCDF.
if (save_plots) pdf(sprintf('%s/tree_metrics_max_nonleaf_degree.pdf', PLOTDIR), width=1.68, height=1.2, pointsize=6,
                    family='Helvetica', useDingbats=FALSE)
par(mar=c(3.4, 3.4, 1.1, 0.6))
plot(sort(metric), rev(1:sum(is.finite(metric)))/sum(is.finite(metric)), log='xy', type='l',
     xlim=c(1,200), ylim=c(1e-5,1),
     panel.first=grid(), xlab='', ylab='')
mtext('d', side=1, line=2.4)
mtext(expression(paste('Pr[max internal deg >= d]')), side=2, line=2.4)
par(.default_par)
if (save_plots) dev.off()
```

Average degree per tree.

```{r}
# PDF
metric <- data$degree_mean
h <- hist(metric, breaks=(floor(min(metric, na.rm=TRUE)-1)):ceiling(max(metric, na.rm=TRUE)), plot=FALSE)
plot(h$breaks[-1], h$counts/sum(h$counts), log='xy', panel.first=grid(),
     xlab='Mean non-leaf degree', ylab='Probability')
# CCDF.
if (save_plots) pdf(sprintf('%s/tree_metrics_avg_nonleaf_degree.pdf', PLOTDIR), width=1.68, height=1.2, pointsize=6,
                    family='Helvetica', useDingbats=FALSE)
par(mar=c(3.4, 3.4, 1.1, 0.6))
plot(sort(metric), rev(1:sum(is.finite(metric)))/sum(is.finite(metric)), log='xy', type='l',
     xlim=c(1,200), ylim=c(1e-5,1),
     panel.first=grid(), xlab='', ylab='')
mtext('d', side=1, line=2.4)
mtext(expression(paste('Pr[avg internal deg >= d]')), side=2, line=2.4)
par(.default_par)
if (save_plots) dev.off()
```

# Wiener index

The Wiener index measures the average path length over all pairs of nodes in the tree. It is maximal for a linear chain and minimal for a tree in which all non-root nodes are children of the root.

```{r}
metric <- data$wiener_index
plot(sort(metric), rev(1:length(metric))/length(metric), log='xy',
     panel.first=grid(), xlab='Wiener index w', ylab='1-F(w)')
```

For a tree consisting of a single click, the Wiener index is 1. Since single-click trees are most likely, a Wiener index of 1 is necessarily most likely. So what if we fix the size $s$ of the tree to some larger value?

The following boxplot shows the distribution of the Wiener index for different values of $s$ (on the $x$-axis). We observe that the Wiener index stays below 5 even as we increase the tree size to $s=30$. This is noteworthy because the theoretical maximum grows linearly and equals roughly $s/3$. This means that the fraction of very long paths (approximately linear chains) decreases with the tree size.

```{r}
range <- 3:30
small_trees <- data[data$size %in% range,]
boxplot(split(small_trees$wiener_index, small_trees$size),
        xlab='Tree size s', ylab='Wiener index')
abline(1, 1/3, col='green', lty=2)
lines(range-min(range)+1, 2*(1-1/range), col='green', lty=3)
legend('topleft', legend=c('Theoretical max', 'Theoretical min'), col='green', lty=c(2,3))
```
